Cyclone: Fault Tolerance Middleware for NVM Clusters

### 2016-05-10 paper/paper.tex

@@

AtomicIncrementSet(uint64_t indices[],
                   uint64_t pre_values[],
                   int count)
{
  bool success = TRUE
  for i = 1 to count
    success =
      success AND
        Prepare(indices[i],
                pre_values[i],
                success)
  for i = 1 to count
    Commit(indices[i], success)       
}
\end{verbatim}
}
\end{minipage}
\end{tabular}
\caption{Counter Set}
\label{fig:counter_set}
\end{figure*}

>>>>

Figure~\ref{fig:counter_set} shows a linearizable implementation executing on
servers. Odd counter values are considered unstable and even values stable.
Clients make RPC calls to servers the trigger the execution of the
appropriate function. Any client can make calls for the get and increment
operations shown on the left hand side. A client retries an operation until it
finds a stable counter value.

A single distinguished client called the co-ordinator runs any atomic increment
set operations on behalf of other clients - shown on the right hand side of
Figure~\ref{fig:counter_set}. We chose to have a single
distinguished co-ordinator to avoid making the example complex by introducting
concurrency.

The operations in Figure~\ref{fig:counter_set} do not comprehend failures.
Cyclone includes a client library to automatically sequence RPC calls at the client
side, retry them, connect to a new quorum leader on a view change and finally
take care of unrealiable networks. At the server side the results of increment,
prepare and commit RPC calls are always persisted and replicated before the
response to the client. For individual get, increment, prepare and commit
operations therefore persistence, replication and failover is transparently
handled.

The AtomicIncrementSet operation presents a special case however. Clients can
fail and therefore a client executing the steps of the operation can fail midway
leaving some counters in an unstable state forever. The solution is the
replicate the distinguished co-ordinator client too. The co-ordinator therefore
becomes both a server, receiving AtomicIncrementSet operations from clients as
well as a client, making individual Prepare and Commit calls to the quorum
leader. The implementation of the AtomicIncrementSet however requires guarantees
beyond that of simple asynchronous replication

\begin{itemize}
  \item Output Commit: AtomicIncrementSet has externally visible side effects at
    other servers when execting. It should therefore be guaranteed to finish
    executing once it starts
  \item Exactly once: Regardless of restarts AtomicIncrementSet should be
    preceived to have executed exactly once by an external observer.
\end{itemize}

The solution is to simply run AtomicIncrementSet in synchronous replication
mode. This guarantees the first property. It also guarantees that exactly once
execution will succced in ``leader'' mode with followers simply remembering the
success status of a transaction. To ensure that due to a failover we do not
prepare or commit a counter more than once, we use exactly once RPCs that
prevent RPC calls from executing more than once. Finally, since only
one AtomicIncrementSet is executed at a time, we include the current step (the
place in the for loop) in RPC calls to the quorum leader and have it reflect the
information back in the response (a cookie of sorts). By querying the results of
the last made call, the co-ordinator can simply fast forward to the correct
index in the for loop during recovery and use the value of the ``success'' field
in the response to update its own local success variable.

@@

AtomicIncrementSet(uint64_t indices[],
                   uint64_t pre_values[],
                   int count)
{
  bool success = TRUE
  for i = 1 to count
    success =
      success AND
        Prepare(indices[i],
                pre_values[i],
                success)
  for i = 1 to count
    Commit(indices[i], success)       
  return success
}
\end{verbatim}
}
\end{minipage}
\end{tabular}
\caption{Counter Set}
\label{fig:counter_set}
\end{figure*}
